\mytitle

### 2020-04-30 paper.tex
Edits for Section 4.2

We measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{The total mean energy is presented in kW.}.
The only method that directly optimizes SS is QP-SS. Hence, for the other methods, we converted solutions
LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in Zhang et al.\ \cite{Zhang2014}. Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
Then, the SS based energy expression can be written as, 
$$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$

@@

In this part, we present the main findings of our empirical evaluation. 
We shall start by showing the best attained solutions by all methods.
Table~\ref{tab:results1} contains the best energy (in kW) attained by the methods
for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. 
The runtime for DA-LS was always less than $10$ seconds.

\begin{table}[t!]

	\begin{tabular}{| c | c | c | c | c | c | c | c |}
		\toprule
		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
		\toprule
		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.10  & 15739.57     \\
		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
		\cline{2-8}
		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}          \\
		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
		&&40   & 21874.04 & 21643.56  & 21905.50 & \textbf{21977.17} & \textbf{21977.17}        \\
		\hline
		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19022.77 \\
		&& 30  & \textbf{27450.99} & 27418.74 & 27442.90 & 27387.06 & 27247.50                     \\
		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34671.62          \\
		\cline{2-8}
		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19172.56 & 19336.33 & 19354.23  & 18448.64            \\
		&&30   & \textbf{27975.55} & 27443.80  & 27752.00 & 27492.11  & 26367.20                      \\
		&&40   & 35604.45 & 35127.32 & \textbf{35665.60}   & 35065.85 & 33451.94 \\
		\bottomrule                   
	\end{tabular}

	\vspace{0.5em}
	\caption{Total Expected Power (kW) after 100 seconds as function of benchmark and method.}\label{tab:results1}
\end{table}

\begin{table}[t!]

	\begin{tabular}{| c | c | c | c | c | c | c | c |}
		\toprule
		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
		\toprule
		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.1  & 15739.57     \\
		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
		\cline{2-8}
		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}          \\
		&&30   & 16570.96 & 16742.66  & 16758.45  & 16755.75 & \textbf{16770.20 }                 \\
		&&40   & 21874.04 & 21832.50  & 21963.81 & \textbf{21977.17} & \textbf{21977.17}        \\
		\hline
		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19213.96 \\
		&& 30  & \textbf{27450.99} & 27442.90 & 27442.90 & 27387.06 & 27389.21                     \\
		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34856.45          \\
		\cline{2-8}
		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19394.58 & 19336.33 & 19354.23  & 19368.86            \\
		&&30   & \textbf{27975.55} & 27785.74  & 27752.00 & 27492.11  & 27633.20                      \\
		&&40   & 35604.45 & 35480.90 & \textbf{35665.60}   & 35065.85 & 35210.78 \\
		\bottomrule                   
	\end{tabular}

	\vspace{0.5em}
	\caption{Total Expected Power (kW) after 3600 seconds as function of benchmark and method.}\label{tab:results2}
\end{table}

We observe that the DA-based solution provides
comparable results to state-of-the-art exact methods. For the harder instances (of $36$ wind directions),
the three quadratic LS-based approaches (DA-LS, QUBO-LS, and QP-LS) dominate the others by finding the best (or equal) solutions in 11 out of 12 instances. For easier instances, 
the linear approach (ILP-LS) finds better solutions \todo{@Chris, Jason: do we have a good explanation for this?}. Table~\ref{tab:results2} shows the same type of results as in Table~\ref{tab:results1}, but now after $3600$ seconds. We see that the DA is still competitive with the other approaches (recall, its solution was found after $10$ seconds),
while QP-LS finds the best solutions in 10 out of the 12 scenarios. For the 6 hard scenarios (12 wind directions), our quadratic solutions
dominate all 6. \todo{@Chris: I am not sure that the first table contributes anything beyond the second table}

We now turn to present the mean relative error over time per approach, i.e., we are presenting $MRE(a,t)$). Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
presents the MRE after $50$ seconds of running the solutions focusing runtime performance, while Figure~\ref{fig:results2} corresponds
to the full run (after $3600$ seconds). The latter provides insights into best performance in terms of energy. The 5 different lines (and their colors) 
correspond to the 5 WFLO solutions. The horizontal axis represents time in seconds.

The results show that on average (across the 12 instances),
the ILP-LS method starts with poor solutions, yet ends up improving over QP-SS and QUBO-LS (after around 1250 seconds).

The two methods that are based on advanced software (QP-LS) and hardware (DA-LS) solutions
outperform the rest both in terms of runtime (best results in short time), and in terms of overall performance.
Specifically, for most instances, QP-LS gives the best solution at 3600 seconds, while the DA-based approach
yields competitive solutions after less than 10 seconds. Furthermore, QP-LS converges to the best solution faster than all other exact methods. In fact, it is inferior only to the approximate DA-LS method, which is geared mainly toward runtime. This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.

>>>>

We measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss})).
%\footnote{The total mean energy is presented in kW.}.
The only method that directly optimizes SS is QP-SS. Hence, for the other methods, we converted solutions
LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in Zhang et al.\ \cite{Zhang2014}. Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
Then, the SS based energy expression can be written as, 
$$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$

@@

%% In this part, we present the main findings of our empirical evaluation. 
%% We shall start by showing the best attained solutions by all methods.
Table~\ref{tab:results1} contains the best energy (in kW) attained by the methods
for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. 
The runtime for DA-LS was always less than $10$ seconds.

\begin{table}[t!]
	\small
	\begin{tabular}{| c | c | c | c | c | c | c | c |}
		\toprule
		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
		\toprule
		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.10  & 15739.57     \\
		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
		\cline{2-8}
		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}          \\
		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
		&&40   & 21874.04 & 21643.56  & 21905.50 & \textbf{21977.17} & \textbf{21977.17}        \\
		\hline
		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19022.77 \\
		&& 30  & \textbf{27450.99} & 27418.74 & 27442.90 & 27387.06 & 27247.50                     \\
		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34671.62          \\
		\cline{2-8}
		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19172.56 & 19336.33 & 19354.23  & 18448.64            \\
		&&30   & \textbf{27975.55} & 27443.80  & 27752.00 & 27492.11  & 26367.20                      \\
		&&40   & 35604.45 & 35127.32 & \textbf{35665.60}   & 35065.85 & 33451.94 \\
		\bottomrule                   
	\end{tabular}

	\vspace{0.5em}
	\caption{Total Expected Power (kW) after 100 seconds for the exact techniques and by 10 seconds for DA-LS.}\label{tab:results1}
\end{table}

\begin{table}[t!]
	\small
	\begin{tabular}{| c | c | c | c | c | c | c | c |}
		\toprule
		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
		\toprule
		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.1  & 15739.57     \\
		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
		\cline{2-8}
		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}          \\
		&&30   & 16570.96 & 16742.66  & 16758.45  & 16755.75 & \textbf{16770.20 }                 \\
		&&40   & 21874.04 & 21832.50  & 21963.81 & \textbf{21977.17} & \textbf{21977.17}        \\
		\hline
		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19213.96 \\
		&& 30  & \textbf{27450.99} & 27442.90 & 27442.90 & 27387.06 & 27389.21                     \\
		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34856.45          \\
		\cline{2-8}
		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19394.58 & 19336.33 & 19354.23  & 19368.86            \\
		&&30   & \textbf{27975.55} & 27785.74  & 27752.00 & 27492.11  & 27633.20                      \\
		&&40   & 35604.45 & 35480.90 & \textbf{35665.60}   & 35065.85 & 35210.78 \\
		\bottomrule                   
	\end{tabular}

	\vspace{0.5em}
	\caption{Total Expected Power (kW) after 3600 seconds and by 10 seconds for DA-LS.}\label{tab:results2}
\end{table}

We observe that the DA-based solution provides
comparable results to state-of-the-art exact methods. For the harder instances (of $36$ wind directions),
the three quadratic LS-based approaches (DA-LS, QUBO-LS, and QP-LS) dominate the others by finding the best (or equal) solutions in 11 out of 12 instances. For easier instances, 
the linear approach (ILP-LS) finds better solutions.
%\todo{@Chris, Jason: do we have a good explanation for this?}. NO
Table~\ref{tab:results2} shows the same type of results as in Table~\ref{tab:results1}, but now after $3600$ seconds. For the exact technique, we can see that the improvement between 100 and 3600 seconds of run-time is marginal. We also observe that the DA is, therefore, still competitive with the other approaches (recall, its solution was found after $10$ seconds),
while QP-LS finds the best solutions in 10 out of the 12 scenarios. Our quadratic approaches dominate all six hard scenarios (12 wind directions). %\todo{@Chris: I am not sure that the first table contributes anything beyond the second table. JCB: ``We include the second table to demonstrate that only minor improvements are achieved by the exact techniques between 100 seconds of run-time.''}

We now turn to present the mean relative error over time per approach. Note that lower values indicate better performance.  Figure~\ref{fig:results1} 
presents the MRE after $50$ seconds, while Figure~\ref{fig:results2} corresponds
to the full run (after $3600$ seconds).
%The latter provides insights into best performance in terms of energy. The 5 different lines (and their colors) 
%correspond to the 5 WFLO solutions. The horizontal axis represents time in seconds.

The results show that on average (across the 12 instances),
the ILP-LS method starts with poor solutions, yet ends up improving over QP-SS and QUBO-LS after around 1250 seconds.

Our proposed quadratic models, QP-LS and DA-LS,
outperform the rest of the techniques both in terms of runtime (best results in short time), and in terms of overall performance.
Specifically, for most instances, QP-LS gives the best solution at 3600 seconds, while the DA-based approach
yields competitive solutions after less than 10 seconds. Furthermore, QP-LS converges to the best solution faster than all other exact methods. In fact, it is inferior only to the approximate DA-LS method.
%This is an encouraging result, especially due to the pessimistic computational complexity of WFLO.
